---
title: "FinalProject"
author: "Jordan Donovan"
date: "April 4, 2017"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Final Project - Breaches

```{r fig1}
library(readr)
library(tidyverse)
require(dplyr)
breaches <- read_csv("~/Downloads/breaches.csv")
View(breaches)
```
We may need to remove Date_of_Breach to tidy the dataset. Really only thing I've seen.


First we will start by exploring the dataset from a breadth first approach. These initial examinations will be where we glean our direction for the following exploration in this document.




Date:

```{r fig2}
breaches %>%
  filter(year > 2009 & year < 2014) %>%
  group_by(year) %>%
  summarise(num_of_cases = n()) %>%
  ggplot(mapping=aes(x=year, y=num_of_cases)) + geom_col()
```


```{r fig3}
breaches %>%
  filter(year > 2009 & year < 2014) %>%
  group_by(year) %>%
  summarise(Individuals_Affected_in_Mil=sum(Individuals_Affected/1000000)) %>%
  ggplot(mapping=aes(x=year, y=Individuals_Affected_in_Mil)) +
    geom_col()
  #geom_smooth()
```


```{r fig4}
#Top 5 types of breaches according to individuals affected by breach
breach_for_pie <- breaches %>%
  filter(year > 2009 & year < 2014) %>%
  group_by(year) %>%
  summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  arrange(desc(Individuals_Affected_in_mil)) %>%
  top_n(5)
slices <- breach_for_pie$Individuals_Affected_in_mil
lbls <- breach_for_pie$year
pct <- round(slices/sum(slices) * 100)
lbls <- paste(lbls, pct)
lbls <- paste(lbls, "%", sep="")
  pie(slices, labels=lbls, col=rainbow(length(lbls)), main="Type of Breach and Individuals Affected in Million")
```

```{r fig4}
require(lubridate)

breaches_date_affected <- breaches %>% filter(!is.na(breach_start) & year < 2014 & year > 2009) %>%
  mutate(month = month(breach_start, label=TRUE)) %>%
  group_by(month) %>%
  summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  ggplot(aes(x=month, y=Individuals_Affected_in_mil)) + geom_col()
breaches_date_affected
```

```{r fig5}
require(lubridate)

make_date_100 <- function(year, month, day) {
  make_date(year, month, day)
}

breaches_by_date_2 <- breaches %>% filter(!is.na(breach_start) & year < 2014 & year > 2009) %>%
  mutate(month = month(breach_start, label = TRUE), day = day(breach_start), breach_start_date = make_date(year, month, day)) %>%
  ggplot(aes(breach_start_date)) + geom_freqpoly(binwidth = month(1) %>% as.numeric())
breaches_by_date_2
```

```{r fig6}
#Plot by month individuals affected
#fit a model to this using MAD or something other than RMS
require(lubridate)
make_date_100 <- function(year, month, day) {
  make_date(year, month, day)
}

breaches_date <- breaches %>% filter(!is.na(breach_start) & year < 2014 & year > 2009) %>%
  mutate(month = month(breach_start, label=TRUE), num_of_occurrences = n()) %>%
  ggplot(aes(x=month, y=num_of_occurrences)) + geom_col(alpha = 1)
breaches_date
```

```{r fig22a}
require(lubridate)

make_date_100 <- function(year, month) {
  make_date(year, month, day)
}

breaches_by_date_2 <- breaches %>% filter(!is.na(breach_start) & year < 2014 & year > 2009) %>%
  mutate(month = month(breach_start, label = TRUE), day = day(breach_start), breach_start_date = make_date(year, month)) %>%
  ggplot(aes(breach_start_date)) + geom_bar()
breaches_by_date_2
```

```{r fig23}
require(lubridate)

make_date_100 <- function(year, month, day) {
  make_date(year, month)
}

breaches_by_date_3 <- breaches %>% filter(!is.na(breach_start) & year < 2014 & year > 2009) %>%
  mutate(month = month(breach_start, label = TRUE), day = day(breach_start), breach_start_date = make_date(year, month)) %>%
  group_by(breach_start_date) %>%
  summarise(Individuals_Affected_in_thou = sum(Individuals_Affected/1000)) %>%
  ggplot(aes(x = breach_start_date, y = Individuals_Affected_in_thou)) + geom_col()
breaches_by_date_3
```

```{r fig24}
require(lubridate)

make_date_100 <- function(year, month, day) {
  make_date(year, month)
}

breaches_by_date4 <- breaches %>% filter(!is.na(breach_start) & year < 2014 & year > 2009) %>%
  mutate(month = month(breach_start, label=TRUE), day=day(breach_start), breach_start_date = make_date(year, month)) %>%
  group_by(breach_start_date) %>%
  summarise(Individuals_Affected_in_thou = sum(Individuals_Affected/1000))

models <- tibble(
    a0 = runif(250, -2000, 4000),
    a1 = runif(250, -5000, 5000)
  )

p2 <- ggplot(breaches_by_date4, aes(breach_start_date, Individuals_Affected_in_thou)) + geom_abline(aes(intercept = a0, slope=a1), data = models, alpha = 1/4) + geom_point()
  p2

  dataset_mod <- lm(Individuals_Affected_in_thou ~breach_start_date, data = breaches_by_date4)
  coef(dataset_mod) %>% str()
  p3 <- summary(dataset_mod)$r.squared * 100
  p3
  
  p4 <- ggplot(breaches_by_date4, aes(breach_start_date, Individuals_Affected_in_thou)) +
  geom_point(size = 2, color = "grey40") +
  geom_abline(intercept = coef(dataset_mod)[1], slope = coef(dataset_mod)[2], color = "blue")
  p4
  
  #grid <- breaches_by_date4 %>%
    #add_predictions(dataset_mod)
  #print(grid)

```

```{r fig24}
require(lubridate)

make_date_100 <- function(year, month, day) {
  make_date(year, month)
}

breaches_by_date5 <- breaches %>% filter(!is.na(breach_start) & year < 2014 & year > 2009) %>%
  mutate(month = month(breach_start, label=TRUE), day=day(breach_start), breach_start_date = make_date(year, month)) %>%
  group_by(breach_start_date) %>%
  summarise(num_of_occurrences = n())

models <- tibble(
    a0 = runif(250, -2000, 4000),
    a1 = runif(250, -5000, 5000)
  )

p2 <- ggplot(breaches_by_date5, aes(breach_start_date, num_of_occurrences)) + geom_abline(aes(intercept = a0, slope=a1), data = models, alpha = 1/4) + geom_point()
  p2

  dataset_mod <- lm(num_of_occurrences ~breach_start_date, data = breaches_by_date5)
  coef(dataset_mod) %>% str()
  p3 <- summary(dataset_mod)$r.squared * 100
  p3
  
  p4 <- ggplot(breaches_by_date5, aes(breach_start_date, num_of_occurrences)) +
  geom_point(size = 2, color = "grey40") +
  geom_abline(intercept = coef(dataset_mod)[1], slope = coef(dataset_mod)[2], color = "blue")
  p4
  
  #grid <- breaches_by_date4 %>%
    #add_predictions(dataset_mod)
  #print(grid)

```





Type/Date

```{r fig7}
breaches %>%
  filter(year > 2009 & year < 2014) %>%
  group_by(Type_of_Breach, year) %>%
  summarise(Individuals_Affected_in_mil=sum(Individuals_Affected/1000000)) %>%
  ggplot(mapping=aes(x=year, size=Individuals_Affected_in_mil, y=Type_of_Breach)) +
  geom_point()
```





Type:

```{r fig8}
breaches %>%
  filter(year > 2009 & year < 2014) %>%
  group_by(Type_of_Breach) %>%
  summarise(Individuals_Affected_in_Mil=sum(Individuals_Affected/1000000)) %>%
  ggplot(mapping=aes(x=Type_of_Breach, y=Individuals_Affected_in_Mil)) +
  geom_col()
```



```{r fig6}
#Top 5 types of breaches according to individuals affected by breach
breaches %>%
  group_by(Type_of_Breach) %>%
  summarise(Individuals_Affected_in_mil=sum(Individuals_Affected/1000000)) %>%
  arrange(desc(Individuals_Affected_in_mil)) %>%
  top_n(5) %>%
  ggplot(mapping=aes(x=Type_of_Breach, y=Individuals_Affected_in_mil, fill=Type_of_Breach)) +
  geom_col()
```

```{r fig6a}
#Top 5 types of breaches according to individuals affected by breach
breach_for_pie <- breaches %>%
  group_by(Type_of_Breach) %>%
  summarise(Individuals_Affected_in_mil=sum(Individuals_Affected/1000000)) %>%
  arrange(desc(Individuals_Affected_in_mil)) %>%
  top_n(5)
slices <- breach_for_pie$Individuals_Affected_in_mil
lbls <- breach_for_pie$Type_of_Breach
pct <- round(slices/sum(slices) * 100)
lbls <- paste(lbls, pct)
lbls <- paste(lbls, "%", sep="")
  pie(slices, labels=lbls, col=rainbow(length(lbls)), main="Type of Breach and Individuals Affected")
```

```{r fig6a}
#Top 5 types of breaches according to individuals affected by breach
breach_for_pie <- breaches %>%
  group_by(Type_of_Breach) %>%
  summarise(num_of_occurrences=n()) %>%
  arrange(desc(num_of_occurrences)) %>%
  top_n(5)
slices <- breach_for_pie$num_of_occurrences
lbls <- breach_for_pie$Type_of_Breach
pct <- round(slices/sum(slices) * 100)
lbls <- paste(lbls, pct)
lbls <- paste(lbls, "%", sep="")
  pie(slices, labels=lbls, col=rainbow(length(lbls)), main="Type of Breach and Number of Occurrences")
```


```{r fig7}
#Top 5 types of breaches according to number of occurrences
breaches %>%
  group_by(Type_of_Breach) %>%
  summarise(num_of_occurrences = n()) %>%
  top_n(5) %>%
  ggplot(mapping=aes(x=Type_of_Breach, y=num_of_occurrences, fill=Type_of_Breach)) +
  geom_col()
```
```{r fig18}
breaches %>% filter(Type_of_Breach %in% (top_Breach_Types$Type_of_Breach)) %>%
  group_by(Type_of_Breach) %>%
  #summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  #mutate(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  mutate(num_of_occurrences = n()) %>%
  mutate(Individuals_Affected_in_thou = (Individuals_Affected/1000)) %>%
  ggplot(mapping=aes(x=num_of_occurrences, color=Type_of_Breach, y=Individuals_Affected_in_thou)) +
  stat_summary()
```

```{r fig15}
breaches %>% filter(Type_of_Breach %in% (top_Breach_Types$Type_of_Breach)) %>%
  ggplot(mapping = aes(x=Type_of_Breach, y=Individuals_Affected, color=Type_of_Breach)) + stat_summary()
```

```{r fig16}
breaches %>% filter(Type_of_Breach %in% (top_Breach_Types$Type_of_Breach)) %>%
  group_by(Type_of_Breach) %>%
  summarise(Individuals_Affected_avg = mean(Individuals_Affected)) %>%
  ggplot(mapping = aes(x=Type_of_Breach, y=Individuals_Affected_avg, fill=Type_of_Breach)) + geom_col(size=5)
```

```{r fig17}
breaches %>% filter(Type_of_Breach %in% (top_Breach_Types$Type_of_Breach)) %>%
  group_by(Type_of_Breach) %>%
  summarise(num_of_occurrences = n()) %>%
  ggplot(mapping = aes(x=Type_of_Breach, y=num_of_occurrences, fill=Type_of_Breach)) + geom_col(size = 5)
```

```{r fig19}
dataset <- breaches %>%
  group_by(Type_of_Breach) %>%
  summarise(Individuals_Affected_in_mil=sum(Individuals_Affected/1000000)) %>%
  arrange(desc(Individuals_Affected_in_mil)) %>%
  top_n(5)

  ggplot(data=dataset, mapping=aes(x=reorder(Type_of_Breach, -Individuals_Affected_in_mil), y=Individuals_Affected_in_mil, fill=Type_of_Breach)) +
  geom_col()

dataset

models <- tibble(
    a0 = runif(250, -20, 40),
    a1 = runif(250, -5, 5)
  )

p2 <- ggplot(dataset, aes(reorder(Type_of_Breach, -Individuals_Affected_in_mil), Individuals_Affected_in_mil)) + geom_abline(aes(intercept = a0, slope=a1), data = models, alpha = 1/4) + geom_point()
  p2

  dataset_mod <- lm(reorder(Type_of_Breach, -Individuals_Affected_in_mil) ~Individuals_Affected_in_mil, data = dataset)
  coef(dataset_mod) %>% str()
  p3 <- summary(dataset_mod)$r.squared * 100
  p3
  
  p4 <- ggplot(dataset, aes(reorder(Type_of_Breach, -Individuals_Affected_in_mil), Individuals_Affected_in_mil)) +
  geom_point(size = 2, color = "grey40") +
  geom_abline(intercept = coef(dataset_mod)[1], slope = coef(dataset_mod)[2], color = "blue")
  p4
  
  #grid <- dataset %>%
    #add_predictions(dataset_mod)
  #print(grid)
```







Type/State:

```{r fig11}
top_Breach_Types <- breaches %>% group_by(Type_of_Breach) %>%
  summarise(Individuals_Affected_in_mil=sum(Individuals_Affected/1000000)) %>%
  arrange(desc(Individuals_Affected_in_mil)) %>%
  top_n(5) %>%
  select(Type_of_Breach)

top_Breach_Types

top_States <- breaches %>%
  group_by(State) %>%
  summarise(Individuals_Affected_in_mil=sum(Individuals_Affected/1000000)) %>%
  arrange(desc(Individuals_Affected_in_mil)) %>%
  top_n(5) %>%
  select(State)
  #group_by(Type_of_Breach, State) %>%
  #summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  #top_n(1) %>%

top_States

both <- breaches %>% filter(Type_of_Breach %in% (top_Breach_Types$Type_of_Breach) & State %in% (top_States$State))

both %>% group_by(Type_of_Breach, State) %>%
  summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  ggplot(mapping=aes(x=State, y=Type_of_Breach, size=Individuals_Affected_in_mil)) +
  geom_point()
```

```{r fig 12}
breaches %>% ggplot(mapping=aes(x=Date_of_Breach, y=State)) + geom_point()
```

```{r fig13}
breaches_by_date <- both %>%
  group_by(Date_of_Breach)
  #summarise(Individuals_Affected_in_mil = sum(Individuals_Affected))
breaches_by_date

ggplot(data=breaches_by_date, mapping=aes(x=Date_of_Breach, y=State)) +
  geom_line() +
  facet_grid(Type_of_Breach ~ .)

##maybe try grouping by date..
```




State:

```{r fig8}
breaches %>%
  group_by(State) %>%
  summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  top_n(15) %>%
  ggplot(mapping = aes(x=State, y=Individuals_Affected_in_mil, fill=State)) +
  geom_col();
```

```{r fig9}
#map(database = 'county')
```




Location:

```{r fig10}
breaches %>%
  group_by(Location_of_Breached_Information) %>%
  summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  top_n(5) %>%
  ggplot(mapping = aes(x=Location_of_Breached_Information, y=Individuals_Affected_in_mil, fill=Location_of_Breached_Information)) +
  geom_col()
```

```{r fig10a}
top_Locations <- breaches %>%
  group_by(Location_of_Breached_Information) %>%
  summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  arrange(desc(Individuals_Affected_in_mil)) %>%
  top_n(5) %>%
  select (Location_of_Breached_Information)

breaches %>% filter(Location_of_Breached_Information %in% top_Locations$Location_of_Breached_Information) %>%
  group_by(Location_of_Breached_Information) %>%
  summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  ggplot(mapping = aes(x=Location_of_Breached_Information, y=Individuals_Affected_in_mil, fill=Location_of_Breached_Information)) +
  geom_col()
```


```{r fig10b}
breaches %>% filter(Location_of_Breached_Information %in% top_Locations$Location_of_Breached_Information) %>%
  group_by(Location_of_Breached_Information) %>%
  #summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  ggplot(mapping = aes(x=Location_of_Breached_Information, y=Individuals_Affected, color=Location_of_Breached_Information)) +
  stat_summary()
```

```{r fig10b}
breaches %>% filter(Location_of_Breached_Information %in% top_Locations$Location_of_Breached_Information) %>%
  group_by(Location_of_Breached_Information) %>%
  mutate(num_of_occurrences = n()) %>%
  mutate(Individuals_Affected_in_thou = (Individuals_Affected/1000)) %>%
  #summarise(Individuals_Affected_in_mil = sum(Individuals_Affected/1000000)) %>%
  ggplot(mapping = aes(x=num_of_occurrences, y=Individuals_Affected, color=Location_of_Breached_Information)) +
  stat_summary()
```



As I begin to analyze this data further, it begins to become apparent that the credibility of this data and its source are not quite at the standard desired.
















